---
title: "Remote Work and Wage Inequality: Correcting Bias in Regression with Generated Binary Labels"
author: "Konrad Kurczynski"
date: "`r Sys.Date()`"
output: 
  md_document:
    variant: gfm
editor_options: 
  markdown: 
    wrap: 72
---
# About this notebook

This notebook estimates the association between working from home and salaries using real-world job postings data [(Hansen et al., 2023)](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=4380734). It illustrates how the functions `ols_bca`, `ols_bcm`, and `one_step` can be used to correct bias from regressing on AI/ML-generated labels.
The notebook reproduces results from Table 1 of [Battaglia, Christensen, Hansen & Sacher (2024)](https://arxiv.org/abs/2402.15585). 

```{r}
library(MLBC)
```

## The Dataset

The package contains a subset of [a larger dataset](https://wfhmap.com) regarding work from home. The sample consists of 16,315 job postings for 2022 and 2023 with "San Diego, CA" recorded as the city and "72" recorded as the NAICS2 industry
code of the advertising firm. 

The dataset contains the following entries: 

1.  `city_name`: city of the job posting

2.  `naics_2022_2` : type of a business

3.  `id`: unique identifier of the job posting

4.  `salary`: salary offered

5.  `wfh_wham`: **ML-generated indicator of whether the job ofers work from home using fine-tuned DistilBERT as in [(Hansen et al., 2023)](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=4380734)**

6.  `soc_2021_2`: Bureau of Labor Statistics Standard Occupational Classification code

7.  `employment_type_name`: indicates whether the position is full-tuime or part-time 


```{r}
SD_data <- MLBC::SD_data
SD_data$salary <- log(SD_data$salary)
head(SD_data)
```

# Estimating the false-positive rate

The variable `wfh_wham` describing whether the job posting offers remote work is not manually collected, but it is imputed via ML methods
using fine-tuned DistilBERT as in [(Hansen et al., 2023)](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=4380734). This classifier has over 99% test accuracy.
Nevertheless as [Battaglia, Christensen, Hansen & Sacher (2024)](https://arxiv.org/abs/2402.15585) document, even high-performance classifiers can lead to large biases in OLS estimates.

The bias correction methods `ols_bca` and `ols_bcm` require estimates of the classifiers false-positive rate. 

We estimate the false positive rate manually. To do so, we took a random sample of size 1000 postings. Of these, 26 had `wfh_wham = 1`. based on reading tese 26 postings, 9 appeared to be misclassified. This means the 
estimated false-positive rate is 0.009. Accordingly, we will implement `ols_bca` and `ols_bcm` with `fpr = 0.009`(the estimated false-positive rate) and `m = 1000` (the sample size used to estimate the false positive rate).

## Results 

We first present results for a simple regression of log alary onto the remote work indicator. We then consider a second specification with fixed effects.

We compare standard OLS estimates and confidence intervals with estimates and confidence intervals using `ols_bcm` which performs a direct bias correction and 
computes bias corrected CIs, and `one_step` which performs maximum likelihod estimation treating the true labels as latent.

### Without fixed effects

We first present OLS estimates: 

```{r}
lm_SD_1 <- ols(salary ~ wfh_wham, data = SD_data)
summary(lm_SD_1)
```

Now using the multiplicative bias correction, with bias corrected CIs:
```{r}
SD_bcm_1 <- ols_bcm(salary ~ wfh_wham, data = SD_data, fpr = 0.009, m = 1000)
summary(SD_bcm_1)
```
Finally, using maximum likelihood:
```{r}
os_SD_1 <- one_step(salary ~ wfh_wham, data = SD_data)
summary(os_SD_1)
```
Note that the one-step estimates here use the default standard normal distribution for the regression errors. These estimates differ from Table 1 of [Battaglia, Christensen, Hansen & Sacher (2024)](https://arxiv.org/abs/2402.15585) which instead uses a Gaussian mixture.

### With fixed effects

We repeat the above now with fixed effects, which are easily generated for the categorical variables `soc_2021_2` amd `employment_type_name`.

First, using OLS:
```{r}
lm_SD_2 <- ols(salary ~ wfh_wham + soc_2021_2 + employment_type_name, data = SD_data)
summary(lm_SD_2)
```

Now using the multiplicative bias correction: 
```{r}
SD_bcm_2 <- ols_bcm(salary ~ wfh_wham + soc_2021_2 + employment_type_name, data = SD_data, fpr = 0.009, m = 1000)
summary(SD_bcm_2)
```

Comparing these results with the OLS results above, we see that the bias corrected CI for the slope coefficient lies to the right of the OLS CI.

Finally, using maximum likelihood:
```{r}
os_SD_2 <- one_step(salary ~ wfh_wham + soc_2021_2 + employment_type_name, data = SD_data)
summary(os_SD_2) 
```
As before, the results for `one_step` diverge from those reported in the "Joint" column of Table 1 in the paper as those were generated using a Gaussian mixture model.


